---
title: "Replicating within plots efffect"
author: "Regan Connell"
date: "5/16/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(nlme)
library(ggplot2)
```

##Exploration 3: Comparison of adding more plots vs. replicating within plots

###Figure 3.1
####Importing simulate_many function from 
```{r}
#Function to simulate 1 split-plot data set that intakse a specified number of plots per oat variety, the true oat and fertilizer effects, as well as the individual random plot effect specified by plot_var (plot variance) and also the overall residual.
simulate_one_data <- function(num_plots, oat_effects, fertilizer_effects, plot_var, residual_sd) {
  
  oat_variety <- c(1, 2)
  treatment_type <- c(1, 0)
  row <- 1
  y <- numeric(0)
  oat <- numeric(0)
  fert <- numeric(0)
  plot <- numeric(0)
  
  for (i in 1:length(oat_effects)) {
    for (pl in 1:num_plots) {
      plot_effect <- rnorm(1, 0, plot_var)
      for (treat in 1:length(fertilizer_effects)) {
        y[row] <- oat_effects[i] + plot_effect + fertilizer_effects[treat] + rnorm(1, 0, residual_sd)
        oat[row] <- oat_variety[i]
        fert[row] <- treatment_type[treat]
        if (i == 1) {
          plot[row] <- pl
        } else {
          plot[row] <- pl + num_plots
        }
        row <- row + 1
      }
    }
  }
  oat <- factor(oat)
  fert <- factor(fert)
  plot <- factor(plot)
  
  
  data <- data.frame(y, oat, fert, plot)
  
  data
}
```

```{r}
#Function to simulate many split-plot data sets with the same parameters as simulate_one_data, with the addition of num_sims to specify the number of simulations to perform. The output will be a list containing each simulations data frames.
simulate_many <- function(num_sims, num_plots, oat_effects, fertilizer_effects, plot_var, residual_sd){
  
  result <- list()
  for (i in 1:num_sims) {
    d <- simulate_one_data(num_plots, oat_effects, fertilizer_effects, plot_var, residual_sd)
    
    result[[i]] <- d
  }
result
}
```

```{r}
#Here is the code to create the data to generate figure 3.1, a figure showing probability that you reject the null hypothesis of no Oat Variety effect, but with plots per Oat Variety on the “x”-axis and different lines for different true Oat variety effects (including no effect).

num_sims31<-90
fert_effects31 <- c(5, 0)
num_plots31 <- 5
num_plots_vector31 <- c(num_plots31, 2*num_plots31, 3*num_plots31,
                        4*num_plots31, 5*num_plots31)
varying_oat_effects31 <- c(0, 1, 2, 3)

plots_per_oat_variety <- numeric(0)
true_oat_effect31 <- numeric(0)
prob_reject31<- numeric(0)
p_vals31 <- numeric(0)


row <-1

for (i in 1:length(num_plots_vector31)) {
  for (j in 1:length(varying_oat_effects31)) {
    for (n in 1:num_sims31) {
      sims31 <- simulate_many(num_sims31, num_plots_vector31[i], c(0, varying_oat_effects31[j]), fert_effects31, 1.5, 1)
      
      p_vals31[n] <- anova(lme(y~oat + fert, random =~1 | plot, data = sims31[[n]]))['oat', 'p-value']
    }
    prob_reject31[row] <- mean(p_vals31 <= 0.05)
    true_oat_effect31[row] <- varying_oat_effects31[j]
    plots_per_oat_variety[row] <- num_plots_vector31[i]
    row <- row +1
  }
}

results31 <- data.frame(prob_reject31, true_oat_effect31, factor(plots_per_oat_variety))
results31

```

```{r}
ggplot(data = results31) +
aes(x = factor(plots_per_oat_variety), y =  prob_reject31, col = factor(true_oat_effect31)) +
geom_point() +
geom_line(aes(group = factor(true_oat_effect31)))
```


###Figure 3.2

```{r}
#This function will simualte one data set of split-plot design with varying replicates.
simulate_one_with_reps <- function(num_reps, num_plots, oat_effects, fert_treatment, plot_var, residual_sd){

  oat_variety <- c(1, 2)
  fertilizer_treatment <- c(1, 0)
  oat <- numeric(0)
  fert <- numeric(0)
  
  plot <- numeric(0)
  y <- numeric(0)
  
  row <- 1
  replicate <- numeric(0)
  for (i in 1:length(oat_variety)){
    for (j in 1:num_plots) {
      plot_effect <- rnorm(1, 0, plot_var)
      for (n in 1:num_reps) {
        for (f in 1:length(fertilizer_treatment)) {
          y[row] <-  plot_effect + rnorm(1, 0, residual_sd) + oat_effects[i] + fert_treatment[f]
          oat[row] <- oat_variety[i]
          fert[row] <- fertilizer_treatment[f]
          plot[row] <- j
          replicate[row] <- n
          row <- row + 1
        }
      }
    }
  }
  
  
  oat <- factor(oat)
  fert <- factor(fert)
  plot <- factor(plot)
  replicate <- factor(replicate)
  data <- data.frame(y, oat, plot, replicate, fert)
  data
}
#Example of a data frame with replicates for sanity check:
simulate_one_with_reps(4, 3, c(1, 3), c(10, 1), 1, 2)
```


```{r}
#This function will simulate many data sets with replicates included using the function simulate_one_with_reps above.
simulate_many_reps <- function(replicates, num_sims, num_plots, oat_effects, fertilizer_effects, plot_var, residual_sd){
  
  result <- list()
  for (i in 1:num_sims) {
    d <- simulate_one_with_reps(replicates, num_plots, oat_effects, fertilizer_effects, plot_var, residual_sd)
    
    result[[i]] <- d
  }
result
}

```


```{r}
#Here is the code to create the data to generate figure 3.2,  a figure showing probability that you reject the null hypothesis of no Oat Variety effect, but with number of Fertilization replicates within each plot on the “x”-axis and different lines for different true Oat variety effects (including no effect).

num_sims32<-90
fert_effects32 <- c(1, 0)
num_plots32 <- 5
varying_fert_replicates32 <- c(0, 5, 10, 15)

varying_oat_effects32 <- c(0, .3, .6, .9)

fert_replicates32 <- numeric(0)
true_oat_effect32 <- numeric(0)
prob_reject32<- numeric(0)
p_vals32 <- numeric(0)


row <-1

for (i in 1:length(varying_oat_effects32)) {
  for (j in 1:length(varying_fert_replicates32)) {
    for (k in 1:num_sims32) {
      sims32 <- simulate_many_reps(varying_fert_replicates32[j], num_sims32, num_plots32, c(0, varying_oat_effects32[i]), fert_effects32, .5, .5)
      
      p_vals32[k] <- anova(lme(y~oat + fert, random =~1 | plot, data = sims32[[k]]))['oat', 'p-value']
    }
    prob_reject32[row] <- mean(p_vals32 <= 0.05)
    true_oat_effect32[row] <- varying_oat_effects32[j]
    fert_replicates32[row] <- varying_fert_replicates32[i]
    row <- row +1
  }
}

results32 <- data.frame(prob_reject32, true_oat_effect32, factor(fert_replicates32))
```


```{r}
ggplot(data = results32) +
aes(x = factor(fert_replicates32), y =  prob_reject32, col = factor(true_oat_effect32)) +
geom_point() +
geom_line(aes(group = factor(true_oat_effect32)))

```




###Figure 3.3

```{r}
#Here is the code to generate data to create a figure showing probability that you reject the null hypothesis of no Fertilization effect, but with plots per Oat Variety on the “x”-axis and different lines for different true Fertilization effects (including no effect).
num_sims33<-90
oat_effects33 <- c(6, 10.5)
num_plots33 <- 5
num_plots_vector33 <- c(num_plots33, 2*num_plots33, 3*num_plots33,
                        4*num_plots33, 5*num_plots33)
varying_fert_effects33 <- c(0, 1, 2, 3)

plots_per_oat_variety <- numeric(0)
true_fert_effect33 <- numeric(0)
prob_reject33<- numeric(0)
p_vals33 <- numeric(0)


row <-1

for (i in 1:length(num_plots_vector33)) {
  for (j in 1:length(varying_fert_effects33)) {
    for (k in 1:num_sims33) {
      sims33 <- simulate_many(num_sims33, num_plots_vector33[i], oat_effects33, c(0, varying_fert_effects33[j]), 1.5, 1)
      
      p_vals33[k] <- anova(lme(y~oat + fert, random =~1 | plot, data = sims33[[k]]))['fert', 'p-value']
    }
    true_fert_effect33[row] <- varying_fert_effects33[j]
    prob_reject33[row] <- mean(p_vals33 <= 0.05)
    plots_per_oat_variety[row] <- num_plots_vector33[i]
    row <- row +1
  }
}

results33 <- data.frame(prob_reject33, true_fert_effect33, factor(plots_per_oat_variety))
results33

```


```{r}
ggplot(data = results33) +
aes(x = factor(plots_per_oat_variety), y =  prob_reject33, col = factor(true_fert_effect33)) +
geom_point() +
geom_line(aes(group = factor(true_fert_effect33)))
```


###Figure 3.4

```{r}
#Here is the code to create the data to generate figure 3.4, a figure showing probability that you reject the null hypothesis of no Fertilization effect, but with number of Fertilization replicates within each plot on the “x”-axis and different lines for different true Fertilization effects (including no effect).

num_sims34<-90
oat_effects34 <- c(1, 0)
num_plots34 <- 5
varying_fert_replicates34 <- c(0, 5, 10, 15)

varying_fert_effects34 <- c(0, .3, .6, .9)

fert_replicates34 <- numeric(0)
true_fert_effect34 <- numeric(0)
prob_reject34<- numeric(0)
p_vals34 <- numeric(0)



row <-1

for (i in 1:length(varying_fert_effects34)) {
  for (j in 1:length(varying_fert_replicates34)) {
    for (k in 1:num_sims34) {
      sims34 <- simulate_many_reps(varying_fert_replicates34[j], num_sims34, num_plots34, oat_effects34, c(0, varying_fert_effects34[i]), .5, .4)
      
      p_vals34[k] <- anova(lme(y~oat + fert, random =~1 | plot, data = sims34[[k]]))['fert', 'p-value']
    }
    prob_reject34[row] <- mean(p_vals34 <= 0.05)
    true_fert_effect34[row] <- varying_fert_effects34[j]
    fert_replicates34[row] <- varying_fert_replicates34[i]
    row <- row +1
  }
}

results34 <- data.frame(prob_reject34, true_fert_effect34, factor(fert_replicates34))
results34

```

```{r}
#Visualize 3.4 data

ggplot(data = results34) +
aes(x = factor(fert_replicates34), y =  prob_reject34, col = factor(true_fert_effect34)) +
geom_point() +
geom_line(aes(group = factor(true_fert_effect34)))

```


3.1 Increasing replicates rather than sample size is more effective at increasing power to detect main effects for oat variety or fertilization.

3.2 It is not possible to achieve 100% power for either kind of effect by increasing the number of data points.










